State of the Practice for Medical Imaging Software

### 2022-07-21 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Further edits to pain points section

With our methodology, we first choose an SC domain (in the current case MI) and
identify a list of about 30 software packages. (For measuring MI we used 29
software packages.)  We approximately measure the qualities of each package by
filling in a grading template. Compared with our previous methodology, the new
methodology also includes repository based metrics, such as the number of files,
number of lines of code, percentage of issues that are closed, etc.  With the
quantitative data in the grading template, we rank the software with the
Analytic Hierarchy Process (AHP) (Details are found in
Section~\ref{ch_background}). After this, as another addition to our previous
methodology, we interview some of the development teams to further understand
the status of their development process.

@@

\citet{PintoEtAl2018} lists some pain points that did not come up in our
conversations with MI developers [CHECK  THESE, after writing all the PPs]:
Cross-platform compatibility, interruptions while coding, scope bloat, lack of
user feedback, hard to collaborate on software projects, and aloneness.
\citet{WieseEtAl2019} repeat some previous pain points and add the following:
dependency management, data handling concerns (like data quality, data
management and data privacy) [Actually mentioned by MI software],
reproducibility, and software scope determination. Although LBM developers did
not mention these pain points, we cannot conclude that they are not relevant for
LBM software development, since we only interviewed four LBM developers for
about an hour each.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} Many interviewees thought lack of time and lack of funding
(discussed next) were their most significant obstacles. Other domains of
research software also experience the lack of time pain point
\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}. Our study of LBM software
\citep{SmithEtAl2022} also highlighted lack of time as a frequently mentioned
pain point.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Incentive and Funding:} As mentioned above, developers felt the pain of having
to attract funding the development and maintenance of their software projects.
For instance, the interviewees from \textit{3D Slicer} and \textit{OHIF} pointed
out that it was more challenging to get fundings for software maintenance as
opposed to research. The interviewee from the \textit{ITK-SNAP} team thought
more fundings was a way to solve the lack of time problem, because they could
hire more dedicated developers. On the other hand, the interviewee from the
\textit{Weasis} team did not feel that fundings could solve the same problem,
since he still would need a lot of time to supervise the project.

Funding challenges have also been noted by others \citep{GewaltigAndCannon2012,
Goble2014, KaterbowAndFeulner2018, SmithEtAl2022}. \citet{WieseEtAl2019}
reported developer pains related to publicity, since publishing norms have
historically made it difficult to get credit for creating software.  As studied
by \citet{HowisonAndBullard2016}, research software (specifically biology
software, but the trend likely applies to other research software domains) is
infrequently cited. \citet{PintoEtAl2018} also mentions the lack of formal
reward system for research software.

An interviewee proposed an idea for increasing funding: Licensing the software
to commercial companies that integrate it into their products.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} Technology pain points for LBM developers include
setting up parallelization and CI. The pain point survey of
\citet{WieseEtAl2019} also highlighted technical-related problems like
dependency management, cross-platform compatibility, CI, hardware issues and
operating system issues. Things mentioned by developers in interviews: Hard to
keep up with changes in OS and libraries, Hard to support multiple OS, and Hard
to support lower-end computers. With respect to balance, developers expressed difficulty balancing between four
factors: cross-platform compatibility, convenience to development \&
maintenance, performance, and security.  The potential and proven solutions are:

% Hard to transfer to new technologies %from Ao

\begin{itemize}
\item Adopting a web-based approach with backend servers, to better support
lower-end computers;
\item Using memory-mapped files to consume less computer memory, to better
support lower-end computers, 
\item Using computing power from the computers GPU for web applications;
\item Increasing funding;
\item Maintaining better documentations to ease the development \& maintenance
processes;
\item Improving performance via more powerful computers, which one interviewee
pointed out has already happened to reduce the balance problem.
\end{itemize}

As the above list shows, web-based applications are perceived to help the
balance problem.  Table~\ref{tab_native_vs_web} shows the teams' choices between
native application and web application. In all the 29 teams on our list, most of
them chose to develop native applications. For the eight teams we interviewed,
three of them were building web applications, and the \textit{MRIcroGL} team was
considering web-based solutions. So we had a good chance to discuss the
differences between the two choices with the interviewees.

@@

The advantage for native applications is higher performance, while web
applications have the advantage of cross-platform compatibility and a simpler
build process.  These web advantages are mirrored by the native application
disadvantages of difficulty with cross-platform compatibility and a complex
build process.  The lower performance disadvantage of web-applications can be
improved with a server backend, but in this case there are disadvantages for
privacy protection and the cost of the servers.

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Ensuring Correctness:} Several developers noted difficulties with
ensuring correctness. They alluded to difficulty with testing the correctness of
large numbers of features, and challenges with automated testing. The TCLB
developer commented that the amount of testing data needed was sometimes
problematic, since free testing services do not offer adequate facilities for
large amounts of data, requiring in-house testing solutions. Other research
software domains point to the following problems with testing: i)
\citet{PintoEtAl2018} mention the problem of insufficient testing; ii)
\citet{HannayEtAl2009} show that more developers think testing is important than
the number that believe they have a sufficient understanding of testing
concepts; and, iii) \citet{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011,
WieseEtAl2019} point to the oracle problem, which occurs in research software
when we do not have a means to judge the correctness of the calculated
solutions. Based on the interviews, the LBM experience seems to overlap with i
and ii; the developers did not allude to problem iii (the oracle problem).  As
discussed in Section~\ref{SecReproducibility}, the oracle problem likely did not
come up because the LBM developers have strategies for verifying their work.

@@

For the pain point category of testing developers identified the lack of access
to real-world datasets.  One pain point in the development process is the lack
of access to real-world datasets for testing. The developers' strategies to
address this are summarized in this section. One threat to correctness is: with
huge datasets for testing, the tests are expensive and time-consuming. Three
interviewees endorsed self tests / automated tests, which may save time for
testing. Some proposed and proven solutions are as follows:

@@

The discussion with the developers focused on understandability issues for two
classes of users: the end users and other developers.  The threats to
understandability for end users include users not understanding how to use
features (\wss{Because the users don't have sufficient background?  Ask Ao.})
\ad{Did you mean medical or medical imaging background? No, it was majorly
caused by the software usability. Based on the context of our interview
conversations, it means that users don't understand how to complete certain
operations or what functions a feature provides. It's due to unclear description,
lack of documentation, poorly designed UI or nonintuitive process},
the team not having a dedicated user experience (UX) designer, some important
indicators are not noticeable (e.g. a progress bar), not all users understand
the purpose of the software, not all users know if the software includes certain
features, not all users understand how to use the command line tool, not all
users understand that the software is a web application. For developers the
threats to understandability include developers not understanding how to deploy
the software (\wss{This sounds like a consequence of a lack of
understandability, rather than threat to it.  Or is the point that developers
lack sufficient background?  Ask Ao}) \ad{Yes, this should be a consequence of a
lack of understandability. I think the direct cause is that the project doesn't
clearly explain how to deploy. The background of the developers wasn't discussed
at that point.}, and the architecture is difficult for new
developers to understand.

>>>>

With our methodology, we first choose an SC domain (in the current case MI) and
identify a list of about 30 software packages. (For measuring MI we used 29
software packages.)  We approximately measure the qualities of each package by
filling in a grading template. Compared with our previous methodology, the new
methodology also includes repository based metrics, such as the number of files,
number of lines of code, percentage of issues that are closed, etc.  With the
quantitative data in the grading template, we rank the software with the
Analytic Hierarchy Process (AHP) (Section~\ref{ch_background} provides details).
After this, as another addition to our previous methodology, we interview some
development teams to further understand the status of their development process.

@@

[CHECK  THESE, after writing all the PPs] \citet{PintoEtAl2018} lists some pain
points that did not come up in our conversations with MI developers :
Cross-platform compatibility, interruptions while coding, scope bloat, lack of
user feedback, hard to collaborate on software projects, and aloneness.
\citet{WieseEtAl2019} repeat some previous pain points and add the following:
dependency management, data handling concerns (like data quality, data
management and data privacy) [Actually mentioned by MI software],
reproducibility, and software scope determination. Although LBM developers did
not mention these pain points, we cannot conclude that they are not relevant for
LBM software development, since we only interviewed four LBM developers for
about an hour each.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} Many interviewees thought lack of time and lack of funding
(discussed next) were their most significant obstacles. Other domains of
research software also experience the lack of time pain point
\citep{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}. Our study of LBM software
\citep{SmithEtAl2022} also highlighted lack of time as a significant pain point.

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Funding:} As mentioned above, developers felt the pain of having to attract
funding to develop and maintain their software. For instance, the interviewees
from \textit{3D Slicer} and \textit{OHIF} said getting funding for software
maintenance is more challenging than for funding for research. The interviewee
from the \textit{ITK-SNAP} team thought more fundings was a way to solve the
lack of time problem, because they could hire more dedicated developers. On the
other hand, the interviewee from the \textit{Weasis} team did not feel that
fundings could solve the same problem, since he still would need time to
supervise the project.

Funding challenges have also been noted by others \citep{GewaltigAndCannon2012,
Goble2014, KaterbowAndFeulner2018, SmithEtAl2022}. Researchers that devote time
to software have the additional challenge that funding agencies do not always
count software when they are judging the academic excellence of the applicant.
\citet{WieseEtAl2019} reported developer pains related to publicity, since
publishing norms have historically made it difficult to get credit for creating
software.  As studied by \citet{HowisonAndBullard2016}, research software
(specifically biology software, but the trend likely applies to other research
software domains) is infrequently cited. \citet{PintoEtAl2018} also mentions the
lack of formal reward system for research software.

An interviewee proposed an idea for increasing funding: Licensing the software
to commercial companies to integrate it into their products.

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} The technology hurdles mentioned by MI develoers
include: hard to keep up with changes in OS and libraries, difficult to transer
to new technologies, hard to support multiple OS, and hard to support lower-end
computers. Developers expressed difficulty balancing between four factors:
cross-platform compatibility, convenience to development and maintenance,
performance, and security.

The pain point survey of \citet{WieseEtAl2019} also highlighted
technical-related problems like dependency management, cross-platform
compatibility, CI, hardware issues and operating system issues. From
\citep{SmithEtAl2022} technology pain points for LBM developers include setting
up parallelization and CI.

The solutions proposed by the MI developers include the following:

\begin{itemize}
\item Adopting a web-based approach with backend servers, to better support
lower-end computers;
\item Using memory-mapped files to consume less computer memory, to better
support lower-end computers, 
\item Using computing power from the computers GPU for web applications;
\item Increasing funding;
\item Maintaining better documentations to ease the development and maintenance
processes;
\item Improving performance via more powerful computers, which one interviewee
pointed out has already happened to reduce the balance problem.
\end{itemize}

As the above list shows, developers perceive web-based applications to help the
balance problem.  Table~\ref{tab_native_vs_web} shows the teams' choices between
native application and web application. In all the 29 teams on our list, most
chose to develop native applications. For the eight teams we interviewed, three
of them were building web applications, and the \textit{MRIcroGL} team was
considering web-based solutions.

@@

The advantage for native applications is higher performance, while web
applications have the advantage of cross-platform compatibility and a simpler
build process.  These web advantages mirror the native disadvantages of
difficulty with cross-platform compatibility and a complex build process.  The
lower performance disadvantage of web applications can be improved with a server
backend, but in this case there are disadvantages for privacy protection and
server costs.

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Ensuring Correctness:} Several developers noted difficulties with
ensuring correctness.

Other research software domains point to the following problems with testing: i)
\citet{PintoEtAl2018} mention the problem of insufficient testing; ii)
\citet{HannayEtAl2009} show that more developers think testing is important than
the number that believe they have a sufficient understanding of testing
concepts; and, iii) \citet{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011,
WieseEtAl2019} point to the oracle problem, which occurs in research software
when we do not have a means to judge the correctness of the calculated
solutions. LBM developers also found difficult with testing including difficulty
with testing the correctness of large numbers of features, and challenges with
automated testing, since free testing services do not offer adequate facilities
for large amounts of data.

@@

For the pain point category of testing developers identified the lack of access
to real-world datasets.  One pain point in the development process is the lack
of access to real-world datasets for testing. One threat to correctness is: with
huge datasets for testing, the tests are expensive and time-consuming. Three
interviewees endorsed self tests / automated tests, which may save time for
testing. Some proposed and proven solutions are as follows:

@@

The discussion with the developers focused on understandability issues for two
classes of users: the end users and other developers.  The threats to
understandability for end users include users not understanding how to use
features (\wss{Because the users don't have sufficient background?  Ask Ao.})
\ad{Did you mean medical or medical imaging background? No, it was majorly
caused by the software usability. Based on the context of our interview
conversations, it means that users don't understand how to complete certain
operations or what functions a feature provides. It's due to unclear
description, lack of documentation, poorly designed UI or nonintuitive process},
the team not having a dedicated user experience (UX) designer, some important
indicators are not noticeable (e.g. a progress bar), not all users understand
the purpose of the software, not all users know if the software includes certain
features, not all users understand how to use the command line tool, not all
users understand that the software is a web application. For developers the
threats to understandability include developers not understanding how to deploy
the software (\wss{This sounds like a consequence of a lack of
understandability, rather than threat to it.  Or is the point that developers
lack sufficient background?  Ask Ao}) \ad{Yes, this should be a consequence of a
lack of understandability. I think the direct cause is that the project doesn't
clearly explain how to deploy. We did not discuss the background of the
developers at that point.}, and the architecture is difficult for new developers
to understand.
